---
title: "p8105_hw2_zl2860"
author: "Zongchao Liu"
date: "9/25/2019"
output: 
  github_document:
    toc: TRUE
    toc_depth: 2


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(knitr)
```

# Problem 1 

## read and clean the Mr. Trash Wheel sheet
```{r message=FALSE}
trash_wheel <- read_excel("./dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  select(dumpster:homes_powered) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))# round()???
```

## read and clean the precipitation data for 2018 sheet
```{r}
prec_2018 <- read_excel("./dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 3, skip = 1)%>%
  janitor::clean_names()%>%
  drop_na(month,total)%>%
  mutate(year = 2018)
```

## read and clean the precipitation data for 2017 sheet
```{r}
prec_2017 <- read_excel("./dataset/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, skip = 1)%>%
  janitor::clean_names() %>%
    drop_na(month,total) %>%
    mutate(year = 2017)
```

## combine precipitation datasets and convert month to a character variable 
```{r}
combined_17_18 <- bind_rows(prec_2017,prec_2018) %>%
  select(month, year, total) %>%
  mutate(month = month.name[month]) 
```
**Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?**

The total precipitation in 2018 was `r sum(pull(prec_2018, total ))` tons and the median number of sports balls in a dumpster in 2017 was `r trash_wheel %>% filter(year == 2017) %>% summarise(median = median(sports_balls))`


# Promblem 2  
## clean the data in "pols-month.csv"  
```{r message = FALSE}
pols_month <- read_csv("./dataset/pols-month.csv") %>%
  separate(mon,
           c("year", "month", "day"),
           remove = TRUE) %>%
  mutate(month = month.abb[as.numeric(month)],
         president = ifelse(prez_dem == "0","gop","dem")) %>%  ###这里会凉,0,1,2???
  select(-day, -prez_dem, -prez_gop)
```

## clean the Data in "snp.csv"
```{r message=FALSE}
snp <- read_csv("./dataset/snp.csv") %>%
  separate(date,
           c("month","day","year")) %>%
  select(year,month,everything(),-day)%>%
  mutate(month = month.abb[as.numeric(month)])
```

## tidy the unemployment data
```{r  message = FALSE }
unemployment <- read_csv("./dataset/unemployment.csv") %>%
  pivot_longer( Jan:Dec,
                names_to = "month",
                values_to = "unemplm_percentage"
  )%>%
  mutate( year = as.character(Year))%>%
  select(-Year) %>%
  select(year, month, everything()) %>%
  drop_na() # after pivot_longer(), 6 rows contain missing values
```

## join the datasets
```{r}
merged_problem2 <- full_join(full_join(pols_month,snp),unemployment)

  
```

# Problem 3 
## Load and tidy the data
```{r message=FALSE}
Popular_Baby_Names <- read_csv("./dataset/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(gender = str_to_lower(gender),
         ethnicity = str_to_lower(recode(ethnicity,
                            "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
                            "BLACK NON HISP" = "BLACK NON HISPANIC" ,
                            "WHITE NON HISP" = "WHITE NON HISPANIC"
                            )),
         childs_first_name = str_to_title(childs_first_name) )%>%
  distinct()
```

##  the rank in popularity of the name “Olivia” as a female baby name over time
```{r}
#The table shows the rank in popularity of the name “Olivia” as a female baby name over time
changes_of_Olivia <- Popular_Baby_Names %>%
  filter(childs_first_name == "Olivia") %>%
  select(ethnicity, rank, year_of_birth) %>%
  pivot_wider(names_from = year_of_birth,
              values_from = rank)%>%
  knitr::kable()
changes_of_Olivia
```

## the most popular name among male children over time
```{r}
# the similar table shows the most popular name among male children over time  
male_rank <- Popular_Baby_Names %>%
  filter(gender == "male") %>%
  group_by(ethnicity)%>%
  filter(rank == min(rank)) %>%
  select(ethnicity, year_of_birth, childs_first_name) %>%
  pivot_wider(names_from = year_of_birth ,
              values_from = childs_first_name)%>%
  knitr::kable()
male_rank
```

## scatter plot
```{r fig.width=8}
# create the dataframe for plotting
male_white_NH_2016 <- Popular_Baby_Names %>%
  filter(gender == "male",
         ethnicity == "white non hispanic",
         year_of_birth == 2016)

#use ggplot to make a scatter plot
male_white_NH_2016%>%
  ggplot(aes(x= rank, y = count)) +geom_point(
   alpha = .4)+labs(x="Rank",y="Count", title = "The Number of Children with a Name Against the Rank in Popularity of That Name")+
  theme_light()+
  theme(plot.title = element_text(hjust = 0.5))
  
```


